Characteristic polynomials of p-adic matrices

### 2017-02-04 charpoly/charpoly.tex
Comment \todo

\subsection*{Previous contributions.} 
\todo{Many ad-hoc for Kedlaya-like point counting algorithms.  Our previous papers.  What else?}

@@

Thanks to the application the framework of differential precision of \cite{caruso-roe-vaccon:14a} in \cite{caruso-roe-vaccon:15a},
we known that
the precision on the characteristic polynomial
$\chi_M$ of a matrix $M \in M_n(\mathbb{Q}_p)$
is determined by the comatrix $com(X I_n -M).$
In this article, we provide:
\begin{enumerate}
\item A factorization of $com(X I_n -M)$ (when $M$ has a cyclic vector).
\item A simple criterion to decide whether there can be a gain in precision
on $\chi_M$ when $M \in M_n(\mathbb{Z}_p).$
\item A $\softO (n^3)$ algorithm with operations in $\Zp$
to compute the optimal precision on each coefficient of $\chi_M$ when $M$
is given with uniform precision on its entries.
\item A $\softO (n^\omega)$ algorithm to compute the 
optimal precision on each coefficient of $\chi_M$
for $M \in M_n (\Qp),$
using divisions.
\end{enumerate}

>>>>

\subsection*{Previous contributions.} 
%\todo{Many ad-hoc for Kedlaya-like point counting algorithms.  Our previous papers.  What else?}

@@

Thanks to the application the framework of differential precision of 
\cite{caruso-roe-vaccon:14a} in \cite{caruso-roe-vaccon:15a}, we known 
that the precision on the characteristic polynomial $\chi_M$ of a matrix 
$M \in M_n(\mathbb{Q}_p)$ is determined by the comatrix $com(X I_n -M).$ 
In this article, we provide:
\begin{enumerate}
\renewcommand{\itemsep}{0pt}
%\item A factorization of $com(X I_n -M)$ (when $M$ has a cyclic vector).
\item A simple criterion to decide whether there can be a gain in precision
on $\chi_M$ when $M \in M_n(\mathbb{Z}_p).$
\item A $\softO (n^3)$ algorithm with operations in $\Zp$ to compute the 
optimal precision on each coefficient of $\chi_M$ when $M$ is given with 
uniform precision on its entries.
\item A $\softO (n^\omega)$ algorithm to compute the 
optimal precision on each coefficient of $\chi_M$
for $M \in M_n (\Qp)$, using divisions.
\item A $\softO (n^\omega)$ algorithm to compute the 
optimal precision on an eigenvalue of $M$.
\end{enumerate}
